---
title: "Some Stuff on Priors"
output: html_notebook
---

To set prior directly
```{r}
library(INLA)

# All priors
names(inla.models()$prior)

# What is available
names(inla.models()$latent$iid$hyper)
# Can set theta as 
inla.models()$latent$iid$hyper$theta

# Define a gamma prior
prec.prior <- list(prec = list(prior = "loggamma", param = c(0.01, 0.01)),
  initial = 4, fixed = FALSE)

# Pass on
f <- y ~ 1 + f(idx, model = "iid", hyper = prec.prior)

```

# Defining new custom priors

Character string with 'table:', followed by the values at which the hyperparameters are evaluated and then followed by the values of the log-density.

```{r}

# to set a Gaussian prior with zero mean and precision 0.001 on θ, the table prior can be defined as follows:
theta <- seq(-100, 100, by = 0.1)
log_dens <- dnorm(theta, 0, sqrt(1 / 0.001), log = TRUE)

gaus.prior <- paste0("table: ", 
  paste(c(theta, log_dens), collapse = " ")
)

# Or alternatively add prior as expression
"expression:
  mean = 0;
  prec = 1000;
  logdens = 0.5 * log(prec) - 0.5 * log (2 * pi);
  logdens = logdens - 0.5 * prec * (theta - mean)^2;
  return(logdens);
"


## Define half-normal prior
# Truncated at zero
HN.prior = "expression:
  tau0 = 0.001;
  sigma = exp(-theta/2);
  log_dens = log(2) - 0.5 * log(2 * pi) + 0.5 * log(tau0);
  log_dens = log_dens - 0.5 * tau0 * sigma^2;
  log_dens = log_dens - log(2) - theta / 2;
  return(log_dens);  
"

## Half-cauchy for sigma
HC.prior  = "expression:
  sigma = exp(-theta/2);
  gamma = 25;
  log_dens = log(2) - log(pi) - log(gamma);
  log_dens = log_dens - log(1 + (sigma / gamma)^2);
  log_dens = log_dens - log(2) - theta / 2;
  return(log_dens);
"

## Half Normal
# Constant is precalculated to save time
HT.prior = "expression:
  sigma = exp(-theta/2);
  nu = 3;
  log_dens = 0 - 0.5 * log(nu * pi) - (-0.1207822);
  log_dens = log_dens - 0.5 * (nu + 1) * log(1 + sigma * sigma);
  log_dens = log_dens - log(2) - theta / 2;
  return(log_dens);
"

## Uniform prior
UN.prior = "expression:
  log_dens = 0 - log(2) - theta / 2;
  return(log_dens);
"


```


## Penalized complexity priors

Priors that penalize departure from a base model

```{r}
library(INLA)
inla.doc()
#?pc.*
```

## Sensitivity analysis with R-INLA

```{r}
# revist practice code from earlier
csize_data <- read.csv (file = "INLA/class_size_data.txt", header = FALSE,
  sep = "", dec = ".")

#Set names
names(csize_data) <- c("clsnr", "pupil", "nlitpre", "nmatpre", "nlitpost",
  "nmatpost", "csize")

#Set NA's
csize_data [csize_data < -1e+29 ] <- NA

#Set class size levels
csize_data$csize <- as.factor(csize_data$csize)
levels(csize_data$csize) <- c("<=19", "20-24", "25-29", ">=30")

# - #
# Priors
prior.list = list(
  default = list(prec = list(prior = "loggamma", param = c(1, 0.00005))),
  half.normal = list(prec = list(prior = HN.prior)),
  half.cauchy = list(prec = list(prior = HC.prior)),
  h.t = list(prec = list(prior = HT.prior)),
  uniform = list(prec = list(prior = UN.prior)),
  pc.prec = list(prec = list(prior = "pc.prec", param = c(5, 0.01)))
) 

# Remove rows with NA's
csize_data2 <- na.omit(csize_data[, -5])

# For prior fit a new model
csize.models <- lapply(prior.list, function(tau.prior) {
  inla(nmatpost ~ 1 + nmatpre + nlitpre + csize +
      f(clsnr, model = "iid", hyper = tau.prior), data = csize_data2,
    control.family = list(hyper = tau.prior))
})

```


## Scaling effects and priors
From the book:

'Briefly, scaling a model means that the generalized variance of the latent effect is one. This scaling will take the values of the scale hyperparameter to a different range, which implies that different priors should be used for scaled and unscaled models. In addition, this also means that precision estimates are comparable between different models and that estimates are less affected by re-scaling covariates in the linear predictor. Furthermore, re-scaling makes the precision invariant to changes in the shape and size of the latent effect.'

```{r}
# Lidar data frompackage SemiPar

library("SemiPar")
data(lidar)

#Data for prediction
xx <- seq(390, 720, by = 5)
# Add data for prediction
new.data <- cbind(range = xx, logratio = NA)
new.data <- rbind(lidar, new.data)

# Set prior on precision
prec.prior <- list(prec = list(param = c(0.001, 0.001)))
# -> note how the sd of the gamma prior has changed. Necessary for the scaling!

#RW1 latent effect
m.rw1 <- inla(logratio ~ 1 + f(range, model = "rw1", constr = FALSE,
    hyper = prec.prior),
  data = new.data, control.predictor = list(compute = TRUE))

#RW1 scaled latent effect
m.rw1.sc <- inla(logratio ~ 1 + f(range, model = "rw1", constr = FALSE,
    scale.model = TRUE, hyper = prec.prior),
  data = new.data, control.predictor = list(compute = TRUE))

#RW2 latent effect
m.rw2 <- inla(logratio ~ 1 + f(range, model = "rw2", constr = FALSE,
    hyper = prec.prior),
  data = new.data, control.predictor = list(compute = TRUE))
#RW2 scaled latent effect
m.rw2.sc <- inla(logratio ~ 1 + f(range, model = "rw2", constr = FALSE,
    scale.model = TRUE,  hyper = prec.prior),
  data = new.data, control.predictor = list(compute = TRUE))
```

