---
title: "7 Spatial effects"
output: html_notebook
---

Different kinds of data available that need to be considered.
More likely in O'Haras examples and inlabru

# Areal data

```{r}
library("spatstat")
library("sp")
library("maptools")

data(bei)

# Create SpatialPoints object
bei.pts <- as(bei, "SpatialPoints")

#Create grid
bei.poly <- as(as.im(bei$window, dimyx=c(5, 10)), "SpatialGridDataFrame")
bei.poly <- as(bei.poly, "SpatialPolygons")

#Number of observations per cell
idx <- over(bei.pts, bei.poly)
tab.idx <- table(idx)
#Add number of trees
d <- data.frame(Ntrees = rep(0, length(bei.poly)))
row.names(d) <- paste0("g", 1:length(bei.poly))
d$Ntrees[as.integer(names(tab.idx))] <- tab.idx

# Finaly create SpatialPolygonsDataFrame
bei.trees <- SpatialPolygonsDataFrame(bei.poly, d)

# Mapping
# INLA considers data sorted by column
idx.mapping <- as.vector(t(matrix(1:50, nrow = 10, ncol = 5)))
bei.trees2 <- bei.trees[idx.mapping, ]
plot(bei.trees2,col = bei.trees$Ntrees)

#Summary statistics of covariates
covs <- lapply(names(bei.extra), function(X) {
  layer <- bei.extra[[X]]
    res <- lapply(1:length(bei.trees2), function(Y) {
      summary(layer[as.owin(bei.trees2[Y, ])])})
    res <- as.data.frame(do.call(rbind, res))

    names(res) <- paste0(X, ".", c("min", "1Q", "2Q", "mean", "3Q", "max"))
    return(res)
    })

covs <- do.call(cbind, covs)

#Add to SPDF
bei.trees2@data <- cbind(bei.trees2@data, covs)

library("spdep")
# Different adjancenies. Queen's or Rook's case
bei.adj.q <- poly2nb(bei.trees2) 
bei.adj.r <- poly2nb(bei.trees2, queen = FALSE)

W.bin <- nb2listw(bei.adj.q, style = "B")
W.rs <- nb2listw(bei.adj.q, style = "W")
W.bin

```

Now fit an INLA model with spatial component

```{r}
library("INLA")

#Log-Poisson regression
m0 <- inla(Ntrees ~ elev.mean + grad.mean, family = "poisson",
  data = as.data.frame(bei.trees2) )

#Log-Poisson regression with random effects
bei.trees2$ID <- 1:length(bei.trees2)
m0.re <- inla(Ntrees ~ elev.mean + grad.mean + f(ID), family = "poisson",
  data = as.data.frame(bei.trees2) )

# As noted above, INLA assumes that the lattice is stored by columns, i.e., a vector with the first column, then followed by the second column and so on. Hence, a proper mapping between the spatial object with the data and the data.frame used in the call to inla is required.

# Spatial with random walk
# RW2d
m0.rw2d <- inla(Ntrees ~ elev.mean + grad.mean +
    f(ID, model = "rw2d", nrow = 5, ncol = 10),
  family = "poisson", data = as.data.frame(bei.trees2),
  control.predictor = list(compute = TRUE),
  control.compute = list(dic = TRUE) )

summary(m0.rw2d)

# Matern2D
m0.m2d <- inla(Ntrees ~ elev.mean + grad.mean +
    f(ID, model = "matern2d", nrow = 5, ncol = 10),
  family = "poisson", data = as.data.frame(bei.trees2),
  control.predictor = list(compute = TRUE) )

summary(m0.m2d)

# Add posterior means to the fitted modules
bei.trees2$RW2D <- m0.rw2d$summary.fitted.values[, "mean"]
bei.trees2$MATERN2D <- m0.m2d$summary.fitted.values[, "mean"]

plot(bei.trees2,col = RColorBrewer::brewer.pal(10,'Spectral'))
```

# Irregular lattice data

```{r}
library("rgdal")
boston.tr <- readOGR(system.file("shapes/boston_tracts.shp",
  package="spData")[1])
boston.adj <- poly2nb(boston.tr)

W.boston <- nb2mat(boston.adj, style = "B") 
W.boston.rs <- nb2mat(boston.adj, style = "W") 

boston.tr$CMEDV2 <- boston.tr$CMEDV
boston.tr$CMEDV2 [boston.tr$CMEDV2 == 50.0] <- NA

# Specification
boston.form  <- log(CMEDV2) ~ CRIM + ZN + INDUS + CHAS + I(NOX^2) + I(RM^2) +
    AGE + log(DIS) + log(RAD) + TAX + PTRATIO + B + log(LSTAT)
boston.tr$ID <- 1:length(boston.tr)

# First a model with spatial random effect
boston.iid <- inla(update(boston.form, . ~. + f(ID, model = "iid")),
  data = as.data.frame(boston.tr),
  control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE),
  control.predictor = list(compute = TRUE)
)
summary(boston.iid)

# Need to transform the posterior marginals

# Use 4 cores to process marginals in parallel
library("parallel")
options(mc.cores = 4)
# Transform marginals and compute posterior mean
#marginals: List of `marginals.fitted.values`from inla model
tmarg <- function(marginals) {
  post.means <- mclapply(marginals, function (marg) {
  # Transform post. marginals
  aux <- inla.tmarginal(exp, marg)
  # Compute posterior mean
  inla.emarginal(function(x) x, aux)
  })

  return(as.vector(unlist(post.means)))
}

# Add posterior means to the SpatialPolygonsDataFrame
boston.tr$IID <- tmarg(boston.iid$marginals.fitted.values)
plot(boston.tr,col=hsv(boston.tr$IID/100))

```

# Latent models in INLA for irregular lattice data.

Besag’s proper spatial model, Besag’s improper spatial model and the one by Besag, York and Mollié, that is a convolution of an intrinsic CAR model and i.i.d. Gaussian model
(besag, besagproper and bym )

```{r}
#Besag's improper
boston.besag <- inla(update(boston.form, . ~. +
    f(ID, model = "besag", graph = W.boston)), 
  data = as.data.frame(boston.tr),
  control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE),
  control.predictor = list(compute = TRUE)
)
boston.tr$BESAG <- tmarg(boston.besag$marginals.fitted.values)

#Besag proper
boston.besagprop <- inla(update(boston.form, . ~. +
    f(ID, model = "besagproper", graph = W.boston)), 
  data = as.data.frame(boston.tr),
  control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE),
  control.predictor = list(compute = TRUE)
)
boston.tr$BESAGPROP <- tmarg(boston.besagprop$marginals.fitted.values)


#BYM
boston.bym <- inla(update(boston.form, . ~. +
    f(ID, model = "bym", graph = W.boston)), 
  data = as.data.frame(boston.tr),
  control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE),
  control.predictor = list(compute = TRUE)
)
boston.tr$BYM <- tmarg(boston.bym$marginals.fitted.values)

plot(boston.tr,col=hsv(boston.tr$BESAG/100))
plot(boston.tr,col=hsv(boston.tr$BESAGPROP/100))
plot(boston.tr,col=hsv(boston.tr$BYM/100))


```


# Geostatistics - Gaussian Processes

Geostatistics deals with the analysis of continuous processes in space. A typical example is the spatial distribution of temperature or pollutants in the air. In this case, the variable of interest is only observed at a finite number of points and statistical methods are required for estimation all over the study region.

Matern covariance
Sigma squared is a general scale parameter. The range of the spatial process is controlled by parameter p. Large values will imply a fast decay in the correlation with distance, which imply a small range spatial process. Small values will indicate a spatial process with a large range. Finally, parameter ν controls smoothness of the spatial process.

```{r}
library("gstat")
library(sp)
library(INLA)
data("meuse")

summary(meuse)

# Make spatial
coordinates(meuse) <- ~x+y
proj4string(meuse) <- CRS("+init=epsg:28992")

#Code from gstat to get the grid
data(meuse.grid)
coordinates(meuse.grid) = ~x+y
proj4string(meuse.grid) <- CRS("+init=epsg:28992")
gridded(meuse.grid) = TRUE


```

## First some kriging

```{r}
# Variogram and fit variogram
vgm <- variogram(log(zinc) ~ dist, meuse)
fit.vgm <- fit.variogram(vgm, vgm("Sph"))

# Krige with square root of predicted variance
krg <- krige(log(zinc) ~ dist, meuse, meuse.grid, model = fit.vgm)

#Add estimates to meuse.grid
meuse.grid$zinc.krg <- krg$var1.pred
meuse.grid$zinc.krg.sd <- sqrt(krg$var1.var)

spplot(meuse.grid)

```

Now model the Meuse dataset with INLA

```{r}
library(maptools)
# First create a boundary of the study region
meuse.bdy <- unionSpatialPolygons(
  as(meuse.grid, "SpatialPolygons"), rep (1, length(meuse.grid))
)

# Define 2d mesh
pts <- meuse.bdy@polygons[[1]]@Polygons[[1]]@coords
mesh <- inla.mesh.2d(loc.domain = pts, max.edge = c(150, 500),
  offset = c(100, 250) )

par(mar = c(0, 0, 0, 0))
plot(mesh, asp = 1, main = "")
lines(pts, col = 3)

# Next create object for the Matern model
meuse.spde <- inla.spde2.matern(mesh = mesh, alpha = 2)
A.meuse <- inla.spde.make.A(mesh = mesh, loc = coordinates(meuse))
s.index <- inla.spde.make.index(name = "spatial.field",
  n.spde = meuse.spde$n.spde)

```

Data passed to ´inla()´ when a SPDE is used needs to be in a particular format. For this, the ´inla.stack´ function is provided. It will take data in different ways, including the SPDE indices, and arrange them conveniently for model fitting. In short, ´inla.stack´ will be a list with the following named elements:

- data: a list with data vectors.

- A: a list of projector matrices.

- effects: a list of effects (e.g., the SPDE index) or predictors (i.e., covariates).

- tag: a character with a label for this group of data.

```{r}
#Create data structure
meuse.stack <- inla.stack(data  = list(zinc = meuse$zinc),
  A = list(A.meuse, 1), # Projector matrix
  effects = list(c(s.index, list(Intercept = 1)),
    list(dist = meuse$dist)),
  tag = "meuse.data") # For indicating that this in the dataset

#Create data structure for prediction
A.pred <- inla.spde.make.A(mesh = mesh, loc = coordinates(meuse.grid))
meuse.stack.pred <- inla.stack(data = list(zinc = NA), # Set reponse to NA
  A = list(A.pred, 1),
  effects = list(c(s.index, list (Intercept = 1)),
    list(dist = meuse.grid$dist)),
  tag = "meuse.pred")

#Join stack
join.stack <- inla.stack(meuse.stack, meuse.stack.pred)

#Fit model
form <- log(zinc) ~ -1 + Intercept + dist + f(spatial.field, model = spde)

# Data passed to the inla function needs to be joined with the definition of the SPDE to fit the spatial model. This is done with function inla.stack.data. Furthermore, control.predictor will need to take the projector matrix for the whole dataset (model fitting and prediction) using function inla.stack.A in join.stack. 

m1 <- inla(form, data = inla.stack.data(join.stack, spde = meuse.spde),
  family = "gaussian",
  control.predictor = list(A = inla.stack.A(join.stack), compute = TRUE),
  control.compute = list(cpo = TRUE, dic = TRUE))

#Get predicted data on grid
index.pred <- inla.stack.index(join.stack, "meuse.pred")$data

meuse.grid$zinc.spde <- m1$summary.fitted.values[index.pred, "mean"]
meuse.grid$zinc.spde.sd <- m1$summary.fitted.values[index.pred, "sd"]

#Compute statistics in terms or range and variance
spde.est <- inla.spde2.result(inla = m1, name = "spatial.field",
  spde = meuse.spde, do.transf = TRUE)

# Kappa
#inla.zmarginal(spde.est$marginals.kappa[[1]])
#Variance
inla.zmarginal(spde.est$marginals.variance.nominal[[1]])
#Range
inla.zmarginal(spde.est$marginals.range.nominal[[1]])

```

# Point patterns - LGCP and co

Previously point patterns were described as discrete characterization of the process, but can equally threat as continious process with a given intensity lambda(chi)

```{r}
library("spatstat")
libary(INLA)
data(clmfires)

#Subsect to 2004 to 2007
clmfires0407 <- clmfires[clmfires$marks$date >= "2004-01-01"]

# Set `urban` instead of `artifgreen`
idx <- which(clmfires.extra$clmcov100$landuse$v %in% c("artifgreen",
  "farm"))
clmfires.extra$clmcov100$landuse$v[idx] <- "urban"

# Convert to factor
clmfires.extra$clmcov100$landuse$v <- 
factor(as.character(clmfires.extra$clmcov100$landuse$v))
# Set right dimension of raster object

dim(clmfires.extra$clmcov100$landuse$v) <- c(100, 100)

#In addition, we will rescale `elevation` (to express it in kilometers) and
#`orientation` (to be in radians) so that fixed effects are better estimated:
clmfires.extra$clmcov100$elevation <- 
  clmfires.extra$clmcov100$elevation / 1000
clmfires.extra$clmcov100$orientation <- 
  clmfires.extra$clmcov100$orientation / (2 * pi)

clmfires0407 <- clmfires0407[clmfires0407$marks$cause == "lightning", ]

# Now create the mesh
clm.bdy <- do.call(cbind, clmfires0407$window$bdry[[1]])
#Define mesh
clm.mesh <- inla.mesh.2d(loc.domain = clm.bdy, max.edge = c(15, 50),
  offset = c(10, 10))

#Points
clm.pts <- as.matrix(coords(clmfires0407))
clm.mesh.pts <- as.matrix(clm.mesh$loc[, 1:2])
allpts <- rbind(clm.pts, clm.mesh.pts)

# Number of vertices in the mesh
nv <- clm.mesh$n
# Number of points in the data
n <- nrow(clm.pts)

# Use penalized complexity prior for the range and standard deviation
# As a prior assumption the probability of the range being higher than 50 is small (i.e., r0=50 and pr=0.9) and that the probability of σ being higher than 5 is also small 
# Create SPDE
clm.spde <- inla.spde2.pcmatern(mesh = clm.mesh, alpha = 2,
  prior.range = c(50, 0.9), # P(range < 50) = 0.9
  prior.sigma = c(1, 0.01) # P(sigma > 10) = 0.01
)

# For the weights use a voroni tesselation
library("deldir")
library("SDraw")

# Voronoi polygons (as SpatialPolygons)
mytiles <- voronoi.polygons(SpatialPoints(clm.mesh$loc[, 1:2]))

# C-LM bounday as SpatialPolygons
clmbdy.sp <- SpatialPolygons(list(Polygons(list(Polygon (clm.bdy)),
  ID = "1"))) 

#Compute weights
require(rgeos)

w <- sapply(1:length(mytiles), function(p) {
  aux <- mytiles[p, ]  

  if(gIntersects(aux, clmbdy.sp) ) {
    return(gArea(gIntersection(aux, clmbdy.sp)))
  } else {
    return(0)
  }
})

# Each polygon weighted by the area
plot(mytiles)

# Values of the covariates at the integration points are required to fit the model. These are obtained from the raster data available in the clmfires.extra dataset.

#Prepare data
y.pp = rep(0:1, c(nv, n))
e.pp = c(w, rep(0, n))

lmat <- inla.spde.make.A(clm.mesh, clm.pts)
imat <- Diagonal(nv, rep(1, nv))

A.pp <-rbind(imat, lmat)

clm.spde.index <- inla.spde.make.index(name = "spatial.field",
  n.spde = clm.spde$n.spde)

# Nearest pixel is used to assign a value
# Covariates
allpts.ppp <- ppp(allpts[, 1], allpts[, 2], owin(xrange = c(-15.87, 411.38), 
  yrange = c(-1.44, 405.19)))

# Assign values of covariates to points using value of nearest pixel
covs100 <- lapply(clmfires.extra$clmcov100, function(X){
  pixels <- nearest.pixel(allpts.ppp$x, allpts.ppp$y, X)
  sapply(1:npoints(allpts.ppp), function(i) {
    X[pixels$row[i], pixels$col[i]]
  })
})
covs100$b0 <- rep(1, nv + n)

# Now for the whole stack, create data structure
clm.stack <- inla.stack(data = list(y = y.pp, e = e.pp),
  A = list(A.pp, 1), 
  effects = list(clm.spde.index, covs100),
  tag = "pp")


# ---------------------------------------- #
# Same structure for prediction
#Data structure for prediction
library("maptools")
sgdf <- as(clmfires.extra$clmcov100$elevation, "SpatialGridDataFrame")
sp.bdy <- as(clmfires$window, "SpatialPolygons")
idx <- over(sgdf, sp.bdy)
spdf <- as(sgdf[!is.na(idx), ], "SpatialPixelsDataFrame")

pts.pred <- coordinates(spdf)
n.pred <- nrow(pts.pred)

#Get covariates (using subsetting operator in spatstat)
ppp.pred <- ppp(pts.pred[, 1], pts.pred[, 2], window = clmfires0407$window)
covs100.pred <- lapply(clmfires.extra$clmcov100, function(X) {
  X[ppp.pred]
})
covs100.pred$b0 <- rep(1, n.pred)

#Prediction points
A.pred <- inla.spde.make.A (mesh = clm.mesh, loc = pts.pred)
clm.stack.pred <- inla.stack(data = list(y = NA),
   A = list(A.pred, 1),
   effects = list(clm.spde.index, covs100.pred), 
   tag = "pred")

#Join data
join.stack <- inla.stack(clm.stack, clm.stack.pred)

# Model fitting will be carried out similarly as in the geostatistics case. However, now the likelihood to be used in a Poisson and the weights associated to the points in the expanded dataset need to be passed using parameter E. The integration strategy is set to "eb" in order to reduce computation time 

# Similar to kernel density smoothing
pp.res0 <- inla(y ~ 1 + 
    f(spatial.field, model = clm.spde), 
  family = "poisson", data = inla.stack.data(join.stack),
  control.predictor = list(A = inla.stack.A(join.stack), compute = TRUE,
    link = 1),
  control.inla = list(int.strategy = "eb"),
  E = inla.stack.data(join.stack)$e)

# And with predictors
pp.res <- inla(y ~ 1 + landuse + elevation + orientation + slope +
    f(spatial.field, model = clm.spde), 
  family = "poisson", data = inla.stack.data(join.stack),
  control.predictor = list(A = inla.stack.A(join.stack), compute = TRUE,
    link = 1), verbose = TRUE,
  control.inla = list(int.strategy = "eb"),
  E = inla.stack.data(join.stack)$e)

summary(pp.res)

#Prediction
idx <- inla.stack.index(join.stack, 'pred')$data
# MOdel with no covariates
spdf$SPDE0 <- pp.res0$summary.fitted.values[idx, "mean"]
#Model with covariates
spdf$SPDE <- pp.res$summary.fitted.values[idx, "mean"]

plot(spdf$SPDE)

# And some statistics, Compute statistics in terms or range and variance
spde.clm.est <- inla.spde2.result(inla = pp.res, name = "spatial.field",
  spde = clm.spde, do.transf = TRUE)

#Variance
inla.zmarginal(spde.clm.est$marginals.variance.nominal[[1]])

```

