---
title: "05 Multiple regression"
output: html_notebook
---

## Multiple regression

Start off with a basic example using a test dataset

```{r}
library(rethinking)
data(WaffleDivorce)
d <- WaffleDivorce

# standardize predictor
d$MedianAgeMarriage.s <- (d$MedianAgeMarriage-mean(d$MedianAgeMarriage))/ sd(d$MedianAgeMarriage)

# fit model
m5.1 <- map(
  alist(
    Divorce ~ dnorm( mu , sigma ) , # Response divorce normal distributed
    mu <- a + bA * MedianAgeMarriage.s , # OLS
    a ~ dnorm( 10 , 10 ) , # intercept normal with 10 sd
    bA ~ dnorm( 0 , 1 ) ,# beta also normal
    sigma ~ dunif( 0 , 10 ) # sd of error uniform
  ) , data = d )


# compute percentile interval of mean
MAM.seq <- seq( from=-3 , to=3.5 , length.out=30 )
mu <- link( m5.1 , data=data.frame(MedianAgeMarriage.s=MAM.seq) )
mu.PI <- apply( mu , 2 , PI )

# plot it all
plot( Divorce ~ MedianAgeMarriage.s , data=d , col=rangi2 )
abline( m5.1 )
shade( mu.PI , MAM.seq )

precis(m5.1)

```
Multiple linear regression on the same dataset but adding median age at marriage

```{r}
d$Marriage.s <- scale(d$Marriage)
m5.3 <- map(
    alist(
      Divorce ~ dnorm( mu , sigma ) ,
      mu <- a + bR*Marriage.s + bA*MedianAgeMarriage.s ,
      a ~ dnorm( 10 , 10 ) ,
      bR ~ dnorm( 0 , 1 ) ,
      bA ~ dnorm( 0 , 1 ) ,
      sigma ~ dunif( 0 , 10 )
    ) ,
  data = d )

precis( m5.3 ) # Posterior mean for marriage rate has +/- for confidence
plot(precis(m5.3))

```
Lets plot some residuals

```{r}
# Residual plot
m5.4 <- map(
      alist(
        Marriage.s ~ dnorm( mu , sigma ) ,
        mu <- a + b*MedianAgeMarriage.s ,
        a ~ dnorm( 0 , 10 ) ,
        b ~ dnorm( 0 , 1 ) ,
        sigma ~ dunif( 0 , 10 )
      ),
    data = d )

# compute expected value at MAP, for each State
mu <- coef(m5.4)['a'] + coef(m5.4)['b']*d$MedianAgeMarriage.s
# compute residual for each State
m.resid <- d$Marriage.s - mu

plot( Marriage.s ~ MedianAgeMarriage.s , d , col=rangi2 )
abline( m5.4 )
# loop over States
for ( i in 1:length(m.resid) ) {
  x <- d$MedianAgeMarriage.s[i] # x location of line segment
  y <- d$Marriage.s[i] # observed endpoint of line segment
  # draw the line segment
  lines( c(x,x) , c(mu[i],y) , lwd=0.5 , col=col.alpha("black",0.7) )
}

```
Counterfactual plot (really just prediction)

```{r}
# prepare new counterfactual data 
A.avg <- mean( d$MedianAgeMarriage.s )
R.seq <- seq( from=-3 , to=3 , length.out=30 )
pred.data <- data.frame(
    Marriage.s=R.seq,
    MedianAgeMarriage.s=A.avg
)
# compute counterfactual mean divorce (mu)
mu <- link( m5.3 , data=pred.data )
mu.mean <- apply( mu , 2 , mean )
mu.PI <- apply( mu , 2 , PI )
# simulate counterfactual divorce outcomes
R.sim <- sim( m5.3 , data=pred.data , n=1e4 )

## ---------- ##
# Essentially holding one variable constant and predict
R.PI <- apply( R.sim , 2 , PI )
# display predictions, hiding raw data with type="n"
plot( Divorce ~ Marriage.s , data=d , type="n" )
mtext( "MedianAgeMarriage.s = 0" )
lines( R.seq , mu.mean )
shade( mu.PI , R.seq )
shade( R.PI , R.seq )

R.avg <- mean( d$Marriage.s )
A.seq <- seq( from=-3 , to=3.5 , length.out=30 )
pred.data2 <- data.frame(
    Marriage.s=R.avg,
    MedianAgeMarriage.s=A.seq
)
mu <- link( m5.3 , data=pred.data2 )
mu.mean <- apply( mu , 2 , mean )
mu.PI <- apply( mu , 2 , PI )
A.sim <- sim( m5.3 , data=pred.data2 , n=1e4 )
A.PI <- apply( A.sim , 2 , PI )

plot( Divorce ~ MedianAgeMarriage.s , data=d , type="n" )
mtext( "Marriage.s = 0" )
lines( A.seq , mu.mean )
shade( mu.PI , A.seq )
shade( A.PI , A.seq )
```

# Posterior predictor plots

Simulare predictions averaging over the posterior

```{r}
# call link without specifying new data
# so it uses original data
mu <- link( m5.3 )
# summarize samples across cases
mu.mean <- apply( mu , 2 , mean )
mu.PI <- apply( mu , 2 , PI )
# simulate observations
# again no new data, so uses original data
divorce.sim <- sim( m5.3 , n=1e4 )
divorce.PI <- apply( divorce.sim , 2 , PI )

plot( mu.mean ~ d$Divorce , col=rangi2 , ylim=range(mu.PI) ,
xlab="Observed divorce" , ylab="Predicted divorce" )
abline( a=0 , b=1 , lty=2 )
for ( i in 1:nrow(d) )
  lines( rep(d$Divorce[i],2) , c(mu.PI[1,i],mu.PI[2,i]) ,col=rangi2 )
```

```{r}
# ------#
# To plot prediction error (residuals)
# compute residuals
divorce.resid <- d$Divorce - mu.mean
# get ordering by divorce rate
o <- order(divorce.resid)
# make the plot
dotchart( divorce.resid[o] , labels=d$Loc[o] , xlim=c(-6,5) , cex=0.6 )
abline( v=0 , col=col.alpha("black",0.2) )
for ( i in 1:nrow(d) ) {
  j <- o[i] # which State in order
  lines( d$Divorce[j]-c(mu.PI[1,j],mu.PI[2,j]) , rep(i,2) )
  points( d$Divorce[j]-c(divorce.PI[1,j],divorce.PI[2,j]) , rep(i,2),
  pch=3 , cex=0.6 , col="gray" )
}
```
<hr>
# Masked relationships

```{r}
data(milk)
d <- milk

# Drop missing data from predictor
dcc <- d[ complete.cases(d) , ]

# Bivariate regression
m5.5 <- map(
    alist(
      kcal.per.g ~ dnorm( mu , sigma ) , # Likelihood
      mu <- a + bn*neocortex.perc , # OLS
      a ~ dnorm( 0 , 100 ) , # intercept
      bn ~ dnorm( 0 , 1 ) , # beta
      sigma ~ dunif( 0 , 1 ) # Sd of likilihood
    ),
data=dcc )

precis(m5.5,digits = 4)

# Change from the smallest percent neocortex percent in the data, 55%, to the largest, 76%, would result in an expected change of only:
coef(m5.5)["bn"] * ( 76 - 55 )

# Quite unplausible given wide range of PI
np.seq <- 0:100 
pred.data <- data.frame( neocortex.perc=np.seq )
mu <- link( m5.5 , data=pred.data , n=1e4 )
mu.mean <- apply( mu , 2 , mean )
mu.PI <- apply( mu , 2 , PI )

plot( kcal.per.g ~ neocortex.perc , data=dcc , col=rangi2 )
lines( np.seq , mu.mean )
lines( np.seq , mu.PI[1,] , lty=2 )
lines( np.seq , mu.PI[2,] , lty=2 )

```

Add log-transformed mass
```{r}
dcc$log.mass <- log(dcc$mass)

m5.6 <- map(
      alist(
          kcal.per.g ~ dnorm( mu , sigma ) ,
          mu <- a + bm*log.mass ,
          a ~ dnorm( 0 , 100 ) ,
          bm ~ dnorm( 0 , 1 ) ,
          sigma ~ dunif( 0 , 1 )
      ) ,
      data=dcc )
precis(m5.6)

m5.7 <- map(
      alist(
      kcal.per.g ~ dnorm( mu , sigma ) ,
      mu <- a + bn*neocortex.perc + bm*log.mass , # Multiple regression with two slopes
      a ~ dnorm( 0 , 100 ) ,
      bn ~ dnorm( 0 , 1 ) ,
      bm ~ dnorm( 0 , 1 ) ,
      sigma ~ dunif( 0 , 1 )
  ) , 
  data = dcc)

m5.7
```

How does predicted energy varies as a function of neocortex percent?

```{r}
mean.log.mass <- mean( log(dcc$mass) )

# 100 prediction values
np.seq <- 0:100
pred.data <- data.frame(
  neocortex.perc=np.seq,
  log.mass=mean.log.mass
)

# link to posterior from model
mu <- link( m5.7 , data=pred.data , n=1e4 )
mu.mean <- apply( mu , 2 , mean )
mu.PI <- apply( mu , 2 , PI )

plot( kcal.per.g ~ neocortex.perc , data=dcc , type="n" )
lines( np.seq , mu.mean )
lines( np.seq , mu.PI[1,] , lty=2 )
lines( np.seq , mu.PI[2,] , lty=2 )

# Both variables cancel each other out
```


# Categorical variables

```{r}

data(Howell1)
d <- Howell1

m5.15 <- map(
      alist(
      height ~ dnorm( mu , sigma ) ,
      mu <- a + bm*male ,
      a ~ dnorm( 178 , 100 ) ,
      bm ~ dnorm( 0 , 10 ) ,
      sigma ~ dunif( 0 , 50 )
      ) ,
  data=d )

precis(m5.15)

post <- extract.samples(m5.15)
mu.male <- post$a + post$bm
PI(mu.male)

# For multiple using an index value
d <- milk
( d$clade_id <- coerce_index(d$clade) )

m5.16 <- map(
  alist(
      kcal.per.g ~ dnorm( mu , sigma ) ,
      mu <- a[clade_id] ,
      a[clade_id] ~ dnorm( 0.6 , 10 ) ,
      sigma ~ dunif( 0 , 10 )
    ), data=d)

precis(m5.16,depth = 2)

```

